My service efforts have been primarily directed towards the CS department
at CU Boulder and the ACM special interest group on programming languages
(SIGPLAN). Representing and serving these communities has been my greatest
honor and privilege.

\vspace*{0.15in}
\mypara{A. Service to the CS Department} I have supported the department in
its educational mission by consistently being available to teach
undergraduate- and graduate-level Programming Languages courses (see
Teaching Statement). I have served on the Graduate Committee for a total of
3 years (Fall 2020-Spring 2022; Fall 2024-Spring 2025). During this time, I
helped the committee make several decisions about student transfers (among
our Master's and PhD programs), course credit transfers (from other
institutions to CU), student fellowships and travel awards, graduate
admissions, and online expansion. The latter refers to launching our
Coursera-based MS-AI program for which I have orchestrated the discussion
within the Graduate Committee and articulated the concerns to the
administration, which led to their speedy resolution and timely approval of
the program. I am currently helping my colleagues Kevin Gifford and Alan
Paradise launch a Coursera specialization in Databases. With respect to
graduate admissions, I have supported the Graduate Committee and the
department on several fronts. Firstly, starting Fall 2020, I have conducted
several research advisory sessions for prospective and new Master's
students. Secondly, I have reviewed hundreds of applications from India
every Spring semester to identify the most promising candidates for
admission to our course-based and research-based Master's programs.
Thirdly, I have helped the committee organize prospective student visit
days by recruiting and organizing student volunteers from CUPLV.  

In addition to my contributions to Graduate Committee, I was part of two
successful faculty searches, which resulted in hiring of Yueqi Chen in 2022
and Paul Krogmeier and Mark Zhao in 2025. The latter two were hired as a
part of Trustworthy and Scalable AI Systems search, which I initiated with
my colleague Evan Chang in Summer 2024. I chaired the search during 2024-25
academic year, coordinating a team of 11 faculty across CUPLV and Systems
groups. We reviewed 387 applications, conducted 18 zoom interviews,
organized campus visits for 4 candidates, interviewed 3, and recommended
offers to 2, both of whom ended up accepting our offers. %For my
%contributions to the department and college's research and teaching
%missions, I was awarded a special recognition by the provost on
%recommendation by the CEAS dean and the CS department chair. 

\vspace*{0.15in}
\mypara{B. Service to ACM SIGPLAN} I am a staunch believer in the potential
of Programming Languages and Formal Methods research and therefore deeply
committed to the advancement of the SIGPLAN research community. The
community has supported me, initially through mentorship programs such as
PLMW and OPLSS, and later by providing meticulous and thoughtful reviews
that improved my work in several cases. It is my pleasure to contribute to
the community by serving on program and organizing committees of SIGPLAN
conferences. Since joining CU, I served on the program committees of PaPoC
2022, 2023, and 2025, PLDI 2023, OOPSLA 2022 and 2024, ATVA 2024, and POPL
2025. I also furnished external reviews to the TOPLAS journal (2023 and
2024) and POPL 2026. For PaPoC 2025, I co-chaired the PC with Davide Frey
from Inria. For OOPSLA 2023 and 2024, I helped the organizing committee by
chairing multiple sessions in the main conference and the SIGPLAN track. In
addition, my students served as student volunteers for flagship SIGPLAN
conferences, including the latest edition of POPL in Denver. In the near
future, besides contributing to PCs, I also plan to mentor PhD students in
Programming Languages via SIGPLAN-M program. 

\vspace*{0.15in}
\mypara{C. Service to Community At Large} As a member of the broader
computing community, I have actively participated in outreach and
engagement activities beyond my departmental and SIGPLAN commitments. I
have served as the CS department's liaison for CU's partnership with
Western Colorado University (WCU) -- a government-designated
hispanic-serving institution. I have organized interactions between CU
faculty and visting WCU students, most of whom are first-generation college
students. Through the CU SPUR program, I have mentored two community
college students from the front range area, introducing them to academic
research in computing. As a registered volunteer for Boulder Valley School
District (BVSD), I mentored two high-school students from Peak-to-Peak
charter school in Lafayette, helping them identify a research project at
the intersection of Generative AI and Security. Additionally, I have
participated in public forums, such as ETH Denver, advocating for
responsible use of blockchain technology.